---
title: 'PHP 2514 HW #5'
author: "Blain Morin"
date: "April 16, 2018"
output: html_document
---

```{r}
library(arm)
library(metRology)
library(knitr)
```

# Question 1: Gellman Hill Chapter 7 #2

According to the textbook, 52% of the United States is female. So, we will run our simulation using this proportion. 

```{r}

### Choose our number of simulations:
n.sims = 1000000

### Initiate a results vector:
totalweight = c()


### Run simulation using a for loop:


for (i in 1:n.sims) {
  
  sex = rbinom(10, 1, .52) ### Determines gender of the ten people (1 = Female)
  
  weight = ifelse(sex == 1, 
                  rnorm(10, mean = 4.96, sd = .2), ### Weight distribution for female 
                  rnorm(10, mean = 5.13, sd = .17)) ### Weight distribution for male
  
  result = sum(exp(weight)) ### Get weight in pounds by exponentiating and add them all together
  
  totalweight[i] = result ### Tally our simulation results
  
}

breaks = sum(totalweight >1750) ### Count the number of cases where the elevator breaks

breaks / n.sims ### Probability of failure 

```

After simulating a million times, we find that the probability that the elevator breaks is .05.

## 2: Gelman Hill Chapter 7 #8

### (a)

How do we use degrees freedom??????????????????

```{r}

### Set number of simulations to 1000
n.sims = 1000

### Set estimates given in the problem
u.cost = 600




```


## 3: Gelman Hill Chapter 8 #1

### (a)

```{r}



x1 = 1:100
x2 = rbinom(100, 1, .5)
y = 3 + .1 * x1 + .5 * x2 + rnorm(100, 0, 1)

model.3a = lm(y ~ x1 + x2)
summary.3a = summary(model.3a)
betas.3a = summary.3a$coefficients[ , 1]
beta.error.3a = summary.3a$coefficients[ , 2]
confidence.68 = cbind(betas.3a - beta.error.3a, betas.3a + beta.error.3a)

confidence.68

```


In this run, the true paramters are all within the 68% intervals.

### (b)

```{r}

### Set number of simulations to 1000
n.sims = 1000

### Initialize a matrix for the parameter estimates
b.cover = matrix(nrow = 1000, ncol = 3)

### Run the simulation
for (i in 1:n.sims) {
  
  b.x1 = 1:100
  b.x2 = rbinom(100, 1, .5)
  y.b = 3 + .1 * b.x1 + .5 * b.x2 + rnorm(100, 0, 1)
  b.model = lm(y.b ~ b.x1 +b.x2)
  summary.3b = summary(b.model)
  est.3b = summary.3b$coefficients[ , 1]
  err.3b = summary.3b$coefficients[ , 2]
  
  ### Set the 68% bounds
  lower.bound = est.3b - err.3b
  upper.bound = est.3b + err.3b
  
  ### These statements test if the true parameters are within the bounds
  int.b = ifelse(3 >= lower.bound[1] && 3 <= upper.bound[1], TRUE, FALSE)
  x1.beta = ifelse(.1 >= lower.bound[2] && .1 <= upper.bound[2], TRUE, FALSE)
  x2.beta = ifelse(.5 >= lower.bound[3] && .5 <= upper.bound[3], TRUE, FALSE)
  
  ### Add the test results to the coverage matrix
  b.cover[i , 1] = int.b
  b.cover[i , 2] = x1.beta
  b.cover[i , 3] = x2.beta
  
}

### The coverage probability is the mean
inter.prob = mean(b.cover[ , 1])
x1.prob = mean(b.cover[ , 2])
x2.prob = mean(b.cover[, 3])

### Report results in a table
probs = cbind(inter.prob, x1.prob, x2.prob)

kable(probs)


```

Thus, the probability that the 68% confidence interval covers the true parameter is .68 for the intercept, .67 for the beta on x1, and .69 for the beta on x2. These are all very close to the expected probability of .68.

### (c)

```{r}

### Set number of simulations to 1000
n.sims = 1000

### Initialize a matrix for the parameter estimates
c.cover = matrix(nrow = 1000, ncol = 3)

### Run the simulation
for (i in 1:n.sims) {
  
  c.x1 = 1:100
  c.x2 = rbinom(100, 1, .5)
  y.c = 3 + .1 * c.x1 + .5 * c.x2 + rt.scaled(100, df = 4, mean = 0, sd = 5)
  c.model = lm(y.c ~ c.x1 +c.x2)
  summary.3c = summary(c.model)
  est.3c = summary.3c$coefficients[ , 1]
  err.3c = summary.3c$coefficients[ , 2]
  
  ### Set the 68% bounds
  lower.bound = est.3c - err.3c
  upper.bound = est.3c + err.3c
  
  ### These statements test if the true parameters are within the bounds
  int.c = ifelse(3 >= lower.bound[1] && 3 <= upper.bound[1], TRUE, FALSE)
  x1.c = ifelse(.1 >= lower.bound[2] && .1 <= upper.bound[2], TRUE, FALSE)
  x2.c = ifelse(.5 >= lower.bound[3] && .5 <= upper.bound[3], TRUE, FALSE)
  
  ### Add the test results to the coverage matrix
  c.cover[i , 1] = int.c
  c.cover[i , 2] = x1.c
  c.cover[i , 3] = x2.c
  
}

### The coverage probability is the mean
inter.prob.c = mean(c.cover[ , 1])
x1.prob.c = mean(c.cover[ , 2])
x2.prob.c = mean(c.cover[, 3])

### Report results in a table
probs.c = cbind(inter.prob.c, x1.prob.c, x2.prob.c)

kable(probs.c)


```

All of the coverage probabilities are near .68. Thus, having errors with a t distribution does not affect our coverge probabilities. 

## Gelman Hill Chapter 11 #4

```{r}

library(readr)
library(lubridate)
cd4 = read_csv("cd4.csv")
cd4$vdate = mdy(cd4$vdate)
str(cd4)

```